---
title: "Mapping Fatalities in States"
author: "Joseph Booker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mapping Fatalities in States}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package provides mapping capability for the accident data file from the [Fatality Analysis Reporting system](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).

Currently, only the accident data file from 2013, 2014, and 2015 is included.

## Direct access to files

To see the path to the installed FARS accident data, use the `make_filename` command with the year as an argument:
```{r}
packaging.r.final::make_filename(2015)
```

## Load accident files into R

The `fars_read(path_name)` function returns a parsed data frames corresponding with the accident reports for the given list of years.

```{r, eval=FALSE}
head(packaging.r.final::fars_read(packaging.r.final::make_filename(2013)), 10)
```
```{r, echo=FALSE, results='asis'}
knitr::kable(head(packaging.r.final::fars_read(packaging.r.final::make_filename(2013)), 10))
```

A full discussion of these columns can be seen on the [FARS Analytical User’s Manual](https://crashstats.nhtsa.dot.gov/Api/Public/ViewPublication/812602).

## Plotting reported accidents per state

To plot the location of accident reports for a given state, the `fars_map_state(state.num, year)` can be used. Note that this function does not vectorize over states or years.

One example is to show the accident reports in California (state number 6) in 2013:
```{r}
packaging.r.final::fars_map_state(6, 2013)
```

Another example is the accident report locations in Florida in 2015:
```{r}
packaging.r.final::fars_map_state(12, 2015)
```

The complete list of state numbers can be found in the [FARS Coding and Validation Manual](https://crashstats.nhtsa.dot.gov/Api/Public/ViewPublication/812559) under the section "State Number – FARS Only."